---
title: "Analysis of Cognition Measure"
author: "Olusoji Oluwafemi Daniel and Owokotomo Olajumoke Evangelina"
date: "18 July 2016"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(markdown)
library(rmarkdown)
library(readr)
library(readxl)
library(dplyr)
library(purrr)
library(tidyr)
library(stringr)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```

### 

```{r readingdata}
cogdata <- read_excel('./Data/UCHE KELECHUKWU (TEST).xls',sheet = 2,col_names = F)
```

### Filtering out and Cleaning Patient Characteristics

```{r}
#removing the first 7 rows
patientDetails <- cogdata %>% slice(1:7)
#drop the 2nd and 3rd rows
patientDetails <- patientDetails %>% select(-c(X2,X3)) %>% data.frame()
#cleaning rownames and turning the dataset around
#and later make it variable names
rownames(patientDetails) <- patientDetails[,1]
patientDetails <- patientDetails %>% t() %>% data.frame()
#cleaning new rownames
rownames(patientDetails) <- gsub('X','',rownames(patientDetails))
patientDetails <- patientDetails[-1,]
patientDetails <- as.tbl(patientDetails)
#Cleaning the numeric variables
patientDetails <- patientDetails %>% mutate(Age=as.numeric(Age),duration.of.stroke=as.numeric(duration.of.stroke))
#Cleaning other variables
#getting patients number
patientDetails <- patientDetails %>% mutate(Patient.no=as.numeric(gsub('patient','',Patient.no)))
#the variable name sex was treated as a level of the factor sex
patientDetails <- patientDetails %>% mutate(sex = factor(sex,levels=c('F','M')))
#cleaning the group variable which has the same problme has sex
patientDetails <- patientDetails %>% mutate(group = factor(group,levels=c('A','B')))
#cleaning affected.limb variable which has the same problme has sex
patientDetails <- patientDetails %>% mutate(affected.limb = factor(affected.limb,levels=c('lt','rt')))
#cleaning Handedness variable
patientDetails <- patientDetails %>% mutate(Handedness = tolower(Handedness))
patientDetails <- patientDetails %>% mutate(Handedness=gsub('right','rt',Handedness))
patientDetails <- patientDetails %>% mutate(Handedness = factor(Handedness,levels=c('rt')))
#
write.table(patientDetails,'./Data/PatientDetails.txt',row.names = F)
```

### Filtering and Cleaning Measured Variables

```{r}
#other measurements aside msi strength of each group
ms <- cogdata %>% slice(which(cogdata$X1!='msi strenght'|is.na(cogdata$X1))) %>% slice(c(1,8:n()))
#renaming spacity appropriately
ms$X1[which(ms$X1=='spasticity')] <- c('wristspasticity','elbowspasticity','shoulderspasticity')
#dealing with the first column
filling <- function(c){
      x <- c(which(!is.na(c)),length(c))
      a <- numeric(length(c))
      a[x] <- c[x]
      a[x[length(x)]] <- c[x[length(x) - 1]]
    for(i in 1:length(x)){
    if(i < length(x)){ 
        j <- i+1
        a[(x[i]+1):(x[j]-1)] <- c[x[i]]
      }
    }
return(a)
}
ms$X1 <- filling(ms$X1)
ms$X2 <- filling(ms$X2)
ms$X1[2] <- 'wrist'
```

## Filtering Out MSI Strength (Non Longitudnal)

```{r}
MSI <- cogdata %>% filter(str_detect(X1,'msi strenght')) %>% select(-c(X2,X3)) %>% data.frame()
MSI[,1] <- c('wrist_msi_strength','elbow_msi_strength','shoulder_msi_strength')
rownames(MSI) <- MSI[,1]
MSI <- MSI %>% t() %>% data.frame()
rownames(MSI) <- gsub('X','',rownames(MSI))
MSI <- MSI %>% slice(-1)
MSI <- sapply(MSI,as.numeric) %>% data.frame()
```

### Joining MSI and Patient Characteristics (MSI Dataset for Analysis)

```{r}
MSID <- bind_cols(patientDetails,MSI)
MSIDdata <- MSID %>% gather(key='msistrength',value='measure',-Patient.no,-sex,-group,-affected.limb,-Age,-duration.of.stroke,-Handedness)
MSIDdata <- MSIDdata %>% mutate(msistrength=gsub('_msi_strength','',msistrength))
```


## Subsetting and Cleaning Wrist (Repeated Measure)

```{r}
##subsetting and cleaning the wrist data
wrist <- ms %>% filter(X1=='wrist' | X1=='Patient no') 
#join the first 3 columns together to form unique names for the variables
wrist <- wrist %>% unite(X1,X1:X3) %>% data.frame()
#set the row names to be equal the first column
rownames(wrist) <- as.character(wrist[,1])
#transpose the dataframe and convert back to dataframe
wrist <- wrist %>% t() %>% data.frame()
rownames(wrist) <- 1:39
#names(wrist) <- c('Patient.no','wrist_fl_active', 'wrist_fl_passive', 'wrist_ext_active','wrist_ext_passive')
wrist <- wrist[-1,]
wrist$Patient.no <- patientDetails$Patient.no
wrist <- wrist[,-1]
wrist <- wrist %>% sapply(X=.,as.numeric) %>% data.frame()
wristlong <- wrist %>% gather(key='bodyparts',value='measure',starts_with('wrist'),-Patient.no)
wristlong <- wristlong %>% separate(bodyparts,c('bodypart','flext','passact'),sep='_')
```

### Merging the Wrist and Patient Characteristics (Dataset for Analysis)

```{r}
wristdata <- merge(patientDetails,wristlong,by='Patient.no')
```

## Subsetting and Cleaning Elbow (Repeated Measure)

```{r}
##subsetting and cleaning the wrist data
elbow <- ms %>% filter(X1=='elbow' | X1=='Patient no') 
#join the first 3 columns together to form unique names for the variables
elbow <- elbow %>% unite(X1,X1:X3) %>% data.frame()
#set the row names to be equal the first column
rownames(elbow) <- as.character(elbow[,1])
#transpose the dataframe and convert back to dataframe
elbow <- elbow %>% t() %>% data.frame()
rownames(elbow) <- 1:39
elbow <- elbow[-1,]
elbow$Patient.no <- patientDetails$Patient.no
elbow <- elbow[,-1]
elbow <- elbow %>% sapply(X=.,as.numeric) %>% data.frame()
elbowlong <- elbow %>% gather(key='bodyparts',value='measure',starts_with('elbow'),-Patient.no)
elbowlong <- elbowlong %>% separate(bodyparts,c('bodypart','flext','passact'),sep='_')

```

## Subsetting and Cleaning Wrist Spasticity

```{r}
##subsetting and cleaning the wrist spasticity data
wristspas <- ms %>% filter(X1=='wristspasticity' | X1=='Patient no') 
#join the first 3 columns together to form unique names for the variables
wristspas <- wristspas %>% unite(X1,X1:X3) %>% data.frame()
#set the row names to be equal the first column
rownames(wristspas) <- as.character(wristspas[,1])
#transpose the dataframe and convert back to dataframe
wristspas <- wristspas %>% t() %>% data.frame()
#cleaning the rownames
rownames(wristspas) <- 1:39
#removing the first row (same as the variable names)
wristspas <- wristspas[-1,]
#removing the first colums (same as patient numbers)
wristspas <- wristspas[,-1]
#correcting the patient numbers
wristspas$Patient.no <- patientDetails$Patient.no
#converting the variables appropriately to numeric
wristspas <- wristspas %>% sapply(X=.,as.numeric) %>% data.frame()
#turning the wide format above to long format
wristspaslong <- wristspas %>% gather(key='bodyparts',value='measure',starts_with('wrist'),-Patient.no)
#separating the variables appropriately
wristspaslong <- wristspaslong %>% separate(bodyparts,c('spasticity','day','prepost'),sep='_')
#cleaning the day variable
wristspaslong <- wristspaslong %>% mutate(day=as.numeric(gsub('day.','',day)))
```

### Merging the Wrist Spasticity and Patient Characteristics (Dataset for Analysis)

```{r}
wristspasdata <- merge(patientDetails,wristspaslong,by='Patient.no')
```

## Subsetting and Cleaning Wrist Spasticity